Random Forests for Automatic Matching Bullet-Lands

### 2015-12-20 imaging-paper.Rnw
centering all pics

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{\linewidth}
\caption{View from the side.\label{fig:sidex3p}}{%
      \includegraphics[width=\linewidth]{images/sidex3p.png}
    }
\end{subfigure}    
\begin{subfigure}[b]{\linewidth}
    \caption{View from the top.\label{fig:topx3p}}{%
    \includegraphics[width=\linewidth]{images/topx3p.png}
    }
\end{subfigure}
\caption{An example of a scan of a bullet land from one groove to the next. }
\end{figure}
Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.

@@

\begin{figure}[hbtp]
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
p1
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35. The orange rectangle shows an example of the smoothing window. Grooves and peaks are detected, if they are not within the same window.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

@@

p3
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=

@@

my.loess$fitted + ylab("Surface measurements")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE>>=
my.loess$resid + ylab("Residuals of loess Fit")
@
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
\end{figure}

\begin{figure}[hbtp]
<<loessfits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
br121.groove <- get_grooves(br121_bullet)
br122.groove <- get_grooves(br122_bullet)
br123.groove <- get_grooves(br123_bullet)
br124.groove <- get_grooves(br124_bullet)
br125.groove <- get_grooves(br125_bullet)
br126.groove <- get_grooves(br126_bullet)

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.7\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections. Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
\end{figure}

\subsection{Varying cross sections}
The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up. This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 450, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,225, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bullet~.) +
    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Rectangles in the back identify a line on one of the bullets.  Matching striation marks are indicated by color filled rectangles and arked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.7\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

\begin{figure}[hbtp]
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.7\\linewidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

\begin{figure}[hbtp]
<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.7\\linewidth', warning=FALSE>>=

@@

\section{Evaluation}
\hh{In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
}
\hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land. For this, we use cross sections 25$\mu m$ apart and require a minimum of 0.9 for the cross correlation. Four lands from different bullets are flagged as problematic in this respect. A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface.}
%
\begin{figure}
\begin{subfigure}[t]{.49\textwidth}
\caption{Br6 Bullet 2-1}
%' <<echo=FALSE>>=
%' b1 <- read.x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
%' image(b1$surface.matrix)
%' @
\includegraphics[width=\linewidth]{images/br6-2-1-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}
\caption{Br9 Bullet 2-4}
\includegraphics[width=\linewidth]{images/br9-2-4-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}
\caption{Unkn Bullet B-2}
\includegraphics[width=\linewidth]{images/b-2-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}
\caption{Unkn Bullet Q-4}
\includegraphics[width=\linewidth]{images/q-4-grey.png}
\end{subfigure}
\caption{\label{fig:fourflags}\hh{Images of the four lands that got flagged during the quality assessment. All of them show scratch marks across the striation marks from the barrel. They will be excluded from the remainder of the analysis.}}
\end{figure}
%
\hh{We exclude these four lands from the further matching considerations and 
 run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e. we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unkown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons. Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM). }
 When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land C-3 and all 120 known lands. Two lands show a high CMS. These correspond to the known matches, shown in Figures~\ref{fig:hamby-perfect}b and c. 
 Unfortunately, not all results are as clear cut. 
It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches to get an overview of what we can reasonably expect from an automated match.

>>>>

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{\linewidth}\centering
\caption{View from the side.\label{fig:sidex3p}}{%
      \includegraphics[width=\linewidth]{images/sidex3p.png}
    }
\end{subfigure}    
\begin{subfigure}[b]{\linewidth}\centering
    \caption{View from the top.\label{fig:topx3p}}{%
    \includegraphics[width=\linewidth]{images/topx3p.png}
    }
\end{subfigure}
\caption{An example of a scan of a bullet land from one groove to the next. }
\end{figure}
Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.

@@

\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
p1
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35. The orange rectangle shows an example of the smoothing window. Grooves and peaks are detected, if they are not within the same window.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

@@

p3
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

@@

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=

@@

my.loess$fitted + ylab("Surface measurements")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE>>=
my.loess$resid + ylab("Residuals of loess Fit")
@
\end{subfigure}
\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
\end{figure}

\begin{figure}[hbtp]
  \centering
<<loessfits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
br121.groove <- get_grooves(br121_bullet)
br122.groove <- get_grooves(br122_bullet)
br123.groove <- get_grooves(br123_bullet)
br124.groove <- get_grooves(br124_bullet)
br125.groove <- get_grooves(br125_bullet)
br126.groove <- get_grooves(br126_bullet)

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}\centering
\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}\centering
    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

@@

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.7\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections. Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
\end{figure}

\subsection{Varying cross sections}
The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up. This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 450, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,225, by=25)
lof <- processBullets(paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}\centering
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}\centering
\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bullet~.) +
    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}\centering
    \caption{Rectangles in the back identify a line on one of the bullets.  Matching striation marks are indicated by color filled rectangles and arked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.7\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

\begin{figure}[hbtp]
  \centering
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.7\\linewidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

\begin{figure}[hbtp]
  \centering
<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.7\\linewidth', warning=FALSE>>=

@@

\section{Evaluation}
\hh{In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
}
\hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land. For this, we use cross sections 25$\mu m$ apart and require a minimum of 0.9 for the cross correlation. Four lands from different bullets are flagged as problematic in this respect. A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface.}
%
\begin{figure}
  \centering
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Br6 Bullet 2-1}
%' <<echo=FALSE>>=
%' b1 <- read.x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
%' image(b1$surface.matrix)
%' @
\includegraphics[width=\linewidth]{images/br6-2-1-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Br9 Bullet 2-4}
\includegraphics[width=\linewidth]{images/br9-2-4-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unkn Bullet B-2}
\includegraphics[width=\linewidth]{images/b-2-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unkn Bullet Q-4}
\includegraphics[width=\linewidth]{images/q-4-grey.png}
\end{subfigure}
\caption{\label{fig:fourflags}\hh{Images of the four lands that got flagged during the quality assessment. All of them show scratch marks across the striation marks from the barrel. They will be excluded from the remainder of the analysis.}}
\end{figure}
%
\hh{We exclude these four lands from the further matching considerations and 
 run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e. we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unkown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons. Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM). }
 When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land C-3 and all 120 known lands. Two lands show a high CMS. These correspond to the known matches, shown in Figures~\ref{fig:hamby-perfect}b and c. 
 Unfortunately, not all results are as clear cut. 
It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches to get an overview of what we can reasonably expect from an automated match.

\begin{figure}[hbtp]
\begin{subfigure}[t]{\textwidth}\centering
\caption{Maximal number of CMS between unknown bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\textwidth'>>=
load("data-25-25/unkn9.RData")
cmsdist <- sapply(reslist, function(x) x$maxCMS)
